### 2020-01-23 CalNotes.tex
high dimensional parameter

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\sigma > 0$ and $W_t$ is a standard one-dimensional Brownian motion. The functions $V, p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}
	V(x) = \sum_{i=1}^N \alpha_i V_i(x),
\end{equation}
for coefficients $\alpha_i \in \R$, $i = 1, \ldots, N$, and smooth functions $V_i\colon \R \to \R$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - K V'(X) \dd t + \sqrt{2K\sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In the following, we will denote $A_i \defeq K\alpha_i$. The coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

@@

\begin{theorem}\label{thm:mainTheorem} Let $\widehat A^\epl(T)$ be defined in \eqref{eq:AHatMixed}. Then, if \corr{Assumptions}, 
	\begin{equation}\label{eq:theoremFirst}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}. Moreover, the following central limit theorem holds
	\begin{equation}\label{eq:theoremSecond}
	\lim_{\epl \to 0} \lim_{T \to \infty} \sqrt{T} \left(A - \widehat A^\epl(T) \right) = \Lambda, \quad \text{in law}, 
	\end{equation}
	where $\Lambda \sim \mathcal N(0, \gamma^2)$ and where 
	\begin{equation}
	\gamma = \sqrt{2\sigma} \frac{\sqrt{\E^{\rho^0}[V'(Z)^2]}}{\E^{\rho^0}[V'(Z)V'(X)]}.
	\end{equation}
\end{theorem}

@@

@@

\bibliographystyle{siam}
\bibliography{../../anmc}
\end{document}

>>>>

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\sigma > 0$ and $W_t$ is a standard one-dimensional Brownian motion. The functions $V, p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}
	V(x) = \sum_{i=1}^N \alpha_i V_i(x),
\end{equation}
for coefficients $\alpha_i \in \R$, $i = 1, \ldots, N$, and smooth functions $V_i\colon \R \to \R$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - K V'(X_t^0) \dd t + \sqrt{2K\sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In the following, we will denote $A_i \defeq K\alpha_i$. The coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

@@

\begin{theorem}\label{thm:mainTheorem} Let $\widehat A^\epl(T)$ be defined in \eqref{eq:AHatMixed}. Then, if \corr{Assumptions}, 
	\begin{equation}\label{eq:theoremFirst}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
%   Moreover, the following central limit theorem holds
%	\begin{equation}\label{eq:theoremSecond}
%	\lim_{\epl \to 0} \lim_{T \to \infty} \sqrt{T} \left(A - \widehat A^\epl(T) \right) = \Lambda, \quad \text{in law}, 
%	\end{equation}
%	where $\Lambda \sim \mathcal N(0, \gamma^2)$ and where 
%	\begin{equation}
%	\gamma = \sqrt{2\sigma} \frac{\sqrt{\E^{\rho^0}[V'(Z)^2]}}{\E^{\rho^0}[V'(Z)V'(X)]}.
%	\end{equation}
\end{theorem}

@@

\subsection{Notes on the high-dimensional parameter}

Let us consider the following stochastic differential equation
\label{multiscale_equation_N}
\d X^\varepsilon_t = \left ( - \alpha \cdot V'(X^\varepsilon_t) - \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon_t}{\varepsilon} \right ) \right ) \d t + \sqrt{2\sigma} \d W_t, 
where $\alpha \in \R^N$ and $V \colon \R \to \R^N$, $V(x) = \begin{pmatrix} V_1(x) \dots V_N(x) \end{pmatrix}^\top$ is a vector of potentials. Its homogenized equation is
\d X_t = - A \cdot V'(X_t) \d t + \sqrt{2\Sigma} \d W_t, 
where $A = K\alpha$ and $\Sigma = K \sigma$. We consider the following estimator for the homogenized drift coefficient $A$
\hat{A} = - M^{-1} h,
where
\label{estimators_N}
M = \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \qquad \text{and} \qquad h = \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \dd X^\varepsilon_t.
Note that repeating the same argument as in the case $N = 1$ we have
\label{formula_e}
\delta \sigma \int_{\R} \int_{\R} V'(z) \phi^\varepsilon(x) \psi^\varepsilon(z) \partial_x R^\varepsilon(x,z) \dd x \dd z = \E^{\rho^\varepsilon} [ ((X^\varepsilon)^2 - (Z^\varepsilon)^2) V''(Z^\varepsilon) ],
and
\label{formula_0}
\delta \Sigma \int_{\R} \int_{\R} V'(z) \phi^0(x) \psi^0(z) \partial_x R^0(x,z) \dd x \dd z = \E^{\rho^0} [ (X^2 - Z^2) V''(Z) ].
Inserting equation \eqref{multiscale_equation_N} into \eqref{estimators_N} we have
h = & - \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) (\alpha \cdot V'(X^\varepsilon_t)) \dd t - \frac{1}{T} \int_0^T \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon_t}{\varepsilon} \right ) V'(Z^\varepsilon_t) \dd t + \frac{\sqrt{2\sigma}}{T} \int_0^T V'(Z^\varepsilon_t) \dd W_t \\
= & - \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \right ) \alpha - \frac{1}{T} \int_0^T \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon_t}{\varepsilon} \right ) V'(Z^\varepsilon_t) \dd t + \frac{\sqrt{2\sigma}}{T} \int_0^T V'(Z^\varepsilon_t) \dd W_t,
\hat{A} = \alpha & + \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \right )^{-1} \left ( \frac{1}{T} \int_0^T \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon_t}{\varepsilon} \right ) V'(Z^\varepsilon_t) \dd t \right ) \\
& - \sqrt{2\sigma} \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \right )^{-1} \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \dd W_t \right ).
Passing to the limit as $T$ goes to infinity and applying the ergodic theorem and the strong law of large numbers for martingale we have
\lim_{T \to \infty} - \sqrt{2\sigma} \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \right )^{-1} \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \dd W_t \right ) = 0,
and
\lim_{T \to \infty} \left ( \frac{1}{T} \int_0^T V'(Z^\varepsilon_t) \otimes V'(X^\varepsilon_t) \dd t \right )^{-1} & \left ( \frac{1}{T} \int_0^T \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon_t}{\varepsilon} \right ) V'(Z^\varepsilon_t) \dd t \right ) \\
& = \E^{\rho^\varepsilon} [V'(Z^\varepsilon) \otimes V'(X^\varepsilon)]^{-1} \E^{\rho^\varepsilon} \left [ \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon}{\varepsilon} \right ) V'(Z^\varepsilon) \right ],
\lim_{T \to \infty} \hat{A} = \alpha + \E^{\rho^\varepsilon} [V'(Z^\varepsilon) \otimes V'(X^\varepsilon)]^{-1} \E^{\rho^\varepsilon} \left [ \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon}{\varepsilon} \right ) V'(Z^\varepsilon) \right ].
An integration by parts yields
\E^{\rho^\varepsilon} \left [ \frac{1}{\varepsilon} p' \left ( \frac{X^\varepsilon}{\varepsilon} \right ) V'(Z^\varepsilon) \right ] = & - \E^{\rho^\varepsilon} [V'(Z^\varepsilon) \otimes V'(X^\varepsilon)] \alpha \\
& + \sigma \int_{\R} \int_{\R} V'(z) \phi^\varepsilon(x) \psi^\varepsilon(z) \partial_x R^\varepsilon(x,z) \dd x \dd z,
and applying formula \eqref{formula_e} we have
\lim_{T \to \infty} \hat{A} = \frac{1}{\delta} \E^{\rho^\varepsilon} [V'(Z^\varepsilon) \otimes V'(X^\varepsilon)]^{-1} \E^{\rho^\varepsilon} [ ((X^\varepsilon)^2 - (Z^\varepsilon)^2) V''(Z^\varepsilon) ].
Passing to the limit as $\varepsilon$ goes to zero we obtain
\lim_{\varepsilon \to 0} \lim_{T \to \infty} \hat{A} = \frac{1}{\delta} \E^{\rho^0} [V'(Z) \otimes V'(X)]^{-1} \E^{\rho^0} [ (X^2 - Z^2) V''(Z) ].
Using formula \eqref{formula_0} and integrating by parts we have
\frac{1}{\delta} \E^{\rho^0} [ (X^2 - Z^2) V''(Z) ] & = \Sigma \int_{\R} \int_{\R} V'(z) \phi^0(x) \psi^0(z) \partial_x R^0(x,z) \dd x \dd z \\
& = \E^{\rho^0} [V'(Z) \otimes V'(X)] A,
\lim_{\varepsilon \to 0} \lim_{T \to \infty} \hat{A} = A.

@@

\bibliographystyle{siam}
\bibliography{anmc}
\end{document}
